---
title: "Homework 4"
author: "Yamuna Dhungana"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F,warning=F,echo=F,fig_height=10,fig_width=6,cache = F)
```

## Instructions windows

Answer all questions stated in each problem. Discuss how your results address each question.

Submit your answers as a pdf, typeset (knitted) from an Rmd file. Include the Rmd file in your submission. You can typeset directly to PDF or typeset to Word then save to PDF In either case, both Rmd and PDF are required. If you are having trouble with .rmd, let us know and we will help you. If you knit to Word, check for any LaTeX commands that will not be compatible with Word.

This file can be used as a template for your submission. Please follow the instructions found under "Content/Begin Here" titled **Homework Formatting**. No code should be included in your PDF submission unless explicitly requested. Use the `echo = F` flag to exclude code from the typeset document.

For any question requiring a plot or graph, answer the question first using standard R graphics (See ?graphics). Then provide a equivalent answer using `library(ggplot2)` functions and syntax. You are not required to produce duplicate plots in answers to questions that do not explicitly require graphs, but it is encouraged. 

You can remove the `Instructions` section from your submission.

## Exercises

### Warning: There are only three questions, however they will require more time coding. You may need to review function calling conventions and whether the optional arguments and their default parameters are appropriate.

1. (Ex. 8.1 in HSAUR, modified for clarity) The \textbf{galaxies} data from \textbf{MASS} contains the velocities of 82 galaxies from six well-separated conic sections of space (Postman et al., 1986, Roeder, 1990). The data are intended to shed light on whether or not the observable universe contains 
superclusters of galaxies surrounded by large voids. The evidence for the existence of superclusters would be the multimodality of the distribution of velocities.(8.1 Handbook)

    a) Construct histograms using the following functions: 
    
        -hist() and ggplot()+geom_histogram()
      
        -truehist() and ggplot+geom_histogram() (make sure that the histograms show proportions, not counts.)
      
        -qplot()
        
    Comment on the shape and properties of the variable based on the five plots. Do you notice any 
    sets of observations clustering? (Hint: You can adjust bin number or bin size as you try to 
    determine the properties of the variable, but use the same bin settings between plots in your
    final analysis. You can also overlay the density function or use the rug command.)
    
```{r}
library(MASS)
library(ggplot2)
data(galaxies, package = "MASS")

# Histogram
hist(galaxies, main = "Base R: Velocity of galaxies",
     xlab= "Galaxies",
     ylab= "Velocity",
     col = "#4dff4d")

ggplot() + aes(galaxies)+
  geom_histogram(binwidth = 5000, breaks = c(seq(5000, 35000, 5000)), boundary = NULL,
                 color = "black", fill= "#ffaa80") +
  labs(title = "ggplot: Velocity of galaxies", x = "Galaxies", y = "Velocity")

## true Histogram
truehist(galaxies,
         xlab="Galaxies",
         ylab="Velocity",
         main="Base R: Velocity of galaxies",
         col="#4dff4d")
Bins = 6 
Breaks = seq(5000, 35000, length.out = Bins + 1)
Binwidth = Breaks[2] - Breaks[1]
ggplot() + aes(galaxies) + geom_histogram(aes(y = ..density..), binwidth = Binwidth, 
                                          breaks = c(seq(5000, 35000, 5000)), boundary = NULL, 
                                          fill = "#ffaa80", color = "black", closed = "left") + 
  labs(title = "ggplot: True Histogram showing galaxies")


# Referance :
# https://www.journaldev.com/38367/density-plot-in-r
# https://bradleyboehmke.github.io/tutorials/histograms



## qplot 

qplot(galaxies) +
  labs(title='base R: Histogram showing galaxies (qplot)',
       x='Velocity',
       y='Frequency',
       col="#4dff4d")


# ggplot has discarded qplot 
BINS <- 30
#BINS <- 6
BREAKS <- seq(5000, 35000, length.out = BINS + 1)
BINWIDTH <- BREAKS[2] - BREAKS[1]
ggplot() + aes(galaxies) +
  geom_histogram(bins = BINS, breaks = BREAKS, binwidth = BINWIDTH, boundary = NULL, 
                 fill = "#ffaa80", color = "black") +
  labs(x = "velocity", y= "Frequency", title = 'ggplot: Histogram showing galaxies (qplot)')


# Referance : http://pages.stat.wisc.edu/~st849-1/Rnotes/qplot.pdf
```

## Answers 1(a)

      From the graphs, it appears that all the graphs for hist and true-hist 
      appear to be the same. The only difference from the graphs is the value
      of the y-axis. Histograms give the frequency. Whereas, true-hist gives 
      the probability of the hist.
      We can see the clustering group clustered in the middle. We also can see
      some clusters on either side of the cluster. We can say that there are 
      three clusters, but in the plot, there is an extra one cluster in the 
      middle cluster. Therefore we can assume there ate four clusters in the
      galaxies data.
      
    
    b) Create a new variable \textit{loggalaxies} = $\log$(galaxies). Repeat part a) using 
    the `loggalaxies` variable. Does this affect your interpretation of the graphs?

```{r}
logalaxies <- log(galaxies)

hist(logalaxies, main = "Base R: Velocity of log-galaxies",
     xlab= "log(Galaxies)",
     ylab= "Velocity",
     col = "#4dff4d")

Bins = 7 
Breaks = seq(9, 10.5, length.out = Bins + 1)
Binwidth = Breaks[2] - Breaks[1]
ggplot() + aes(logalaxies)+
  geom_histogram(binwidth = Binwidth, breaks = Breaks, boundary = NULL, color = "black",
                 fill= "#ffaa80") +
  labs(title = "ggplot: Velocity of log-galaxies", x = "log(Galaxies)", y = "Velocity")

truehist(logalaxies,
         xlab="log(Galaxies)",
         ylab="Velocity",
         main="Base R: Velocity of log-galaxies",
         col="#4dff4d")
Bins = 7 
Breaks = seq(9, 10.5, length.out = Bins + 1)
Binwidth = Breaks[2] - Breaks[1]
ggplot() + aes(logalaxies) + geom_histogram(aes(y = ..density..), binwidth = Binwidth, 
                                            breaks = Breaks, boundary = NULL, fill = "#ffaa80",
                                            color = "black", closed = "left") + 
  labs(title = "ggplot: True Histogram showing galaxies")


qplot(logalaxies) +
  labs(title='base R: Histogram showing log-galaxies (qplot)',
       xlab='log(galaxies',
       ylab ='Velocity',
       col="#4dff4d")


Bins = 30 
Breaks = seq(9, 10.5, length.out = Bins + 1)
Binwidth = Breaks[2] - Breaks[1]
ggplot() + aes(logalaxies) +
  geom_histogram(bins = Bins, breaks = Breaks, binwidth = Binwidth, boundary = NULL, fill = "#ffaa80",
                 color = "black") +
  labs(x = "velocity", y= "Frequency", title = 'ggplot: Histogram showing galaxies (qplot)')


# Referance :
# https://www.journaldev.com/38367/density-plot-in-r
# https://bradleyboehmke.github.io/tutorials/histograms
# http://pages.stat.wisc.edu/~st849-1/Rnotes/qplot.pdf

```

## Answer-1(b)

       In this question, we constructed the graph, which is the same as the
       question-a but in the log form. The scale of the values in the log 
       form for the galaxies. A similar description applies to this question
       as well, and we can say that we can see three main clusters with one 
       small cluster in between the concentrated cluster.
      
    
    c) Construct kernel density estimates using two different choices of kernel functions and three 
    choices of bandwidth (one that is too large and “oversmooths,” one that is too small and 
    “undersmooths,” and one that appears appropriate.) Therefore you should have six different 
    kernel density estimates plots (you may combine plots when appropriate to reduce the number 
    of plots made). Discuss your results. You can use the log scale or original scale for the 
    variable, and specify in the plot x-axis which you choose.
    
```{r}
library(KernSmooth)
mydata <- as.data.frame(galaxies)


#plot.new()
# Gassuian with Bandwidh - 5000
# truehist(galaxies,ymax=0.0002,col="green", main="base R: Gaussian Over Smooth with bandwidth=5000")+
# lines(density(galaxies,kernel="gaussian",bw=5000),col="red")
# 
# #ggplot
# Bins = 6 
# Breaks = seq(5000, 35000, length.out = Bins + 1)
# Binwidth = Breaks[2] - Breaks[1]
# ggplot(mydata, aes(x=galaxies)) + 
#   geom_histogram(aes(y=..density..),# Histogram with density instead of count on y-axis
#                  bins = Bins,
#                  binwidth=Binwidth,
#                  colour="black", fill="#ffaa80") +
#   stat_density(kernel='gaussian', bw = 5000, fill = NA, col = "red") +
#   labs(title = "ggplot: Gaussian Over Smooth")  # Overlay with transparent density plot
# 
# 
# ### kernel = Triangular 
# truehist(galaxies,ymax=0.0002,col="green", main="base R: triangular Over Smooth with bandwidth=5000")+
# lines(density(galaxies,kernel="triangular",bw=5000),col="red")
# 
# 
# # ggplot
# Bins = 6 
# Breaks = seq(5000, 35000, length.out = Bins + 1)
# Binwidth = Breaks[2] - Breaks[1]
# ggplot(mydata, aes(x=galaxies)) + 
#   geom_histogram(aes(y=..density..),# Histogram with density instead of count on y-axis
#                  bins = Bins,
#                  binwidth=Binwidth,
#                  colour="black", fill="#ffaa80") +
#   stat_density(kernel='triangular', bw = 5000, fill = NA, col = "red") +
#   labs(title = "ggplot: triangular Over Smooth")  # Overlay with transparent density plot
# 
# 
# 
# 
# #######
# # Gassuian Base R with bw= 800
# truehist(galaxies,ymax=0.0002,col="green", main="base R: Gaussian Over Smooth with bandwidth=800")+
# lines(density(galaxies,kernel="gaussian",bw=800),col="red")
# 
# ## ggplot
# Bins = 6 
# Breaks = seq(5000, 35000, length.out = Bins + 1)
# Binwidth = Breaks[2] - Breaks[1]
# ggplot(mydata, aes(x=galaxies)) + 
#   geom_histogram(aes(y=..density..),# Histogram with density instead of count on y-axis
#                  bins = Bins,
#                  binwidth=Binwidth,
#                  colour="black", fill="#ffaa80") +
#   stat_density(kernel='gaussian', bw = 800, fill = NA, col = "red") +
#   labs(title = "ggplot: Gaussian Over Smooth")  # Overlay with transparent density plot
# 
# ## Kernel = triagular
# 
# truehist(galaxies,ymax=0.0002,col="green", main="base R: triangular Over Smooth with bandwidth=800")+
# lines(density(galaxies, kernel="triangular",bw=800),col="red")
# 
# ## ggplot
# Bins = 6 
# Breaks = seq(5000, 35000, length.out = Bins + 1)
# Binwidth = Breaks[2] - Breaks[1]
# ggplot(mydata, aes(x=galaxies)) + 
#   geom_histogram(aes(y=..density..),# Histogram with density instead of count on y-axis
#                  binwidth=Binwidth,
#                  colour="black", fill="#ffaa80") +
#   stat_density(kernel='triangular', bw = 800, fill = NA, col = "red") +
#   labs(title = "ggplot: triangular Over Smooth")  # Overlay with transparent density plot
# 
# 
# 
# 
# 
# 
# #####
#  truehist(galaxies,ymax=0.0002,col="green", main="base R: Gaussian Over Smooth with bandwidth=100")+
# lines(density(galaxies,kernel="gaussian",bw=100),col="red")
# 
# Bins = 6
# Breaks = seq(5000, 35000, length.out = Bins + 1)
# Binwidth = Breaks[2] - Breaks[1]
# ggplot(mydata, aes(x=galaxies)) +
#   geom_histogram(aes(y=..density..),# Histogram with density instead of count on y-axis
#                  binwidth=Binwidth,
#                  colour="black", fill="#ffaa80") +
#   stat_density(kernel='gaussian', bw = 100, fill = NA, col = "red") +
#   labs(title = "ggplot: Gaussian Over Smooth")  # Overlay with transparent density plot
# 
# ## kernel = Triangular
# 
# truehist(galaxies,ymax=0.0002,col="green", main="base R: Triangular Over Smooth with bandwidth=100")+
# lines(density(galaxies,kernel="triangular",bw=100),col="red")
# 
# Bins = 6
# Breaks = seq(5000, 35000, length.out = Bins + 1)
# Binwidth = Breaks[2] - Breaks[1]
# ggplot(mydata, aes(x=galaxies)) +
#   geom_histogram(aes(y=..density..),# Histogram with density instead of count on y-axis
#                  binwidth=Binwidth,
#                  colour="black", fill="#ffaa80") +
#   stat_density(kernel='triangular', bw = 100, fill = NA, col = "red") +
#   labs(title = "ggplot: Triangular Over Smooth")  # Overlay with transparent density plot
# 

## GGplot

Bins = 6 
Breaks = seq(5000, 35000, length.out = Bins + 1)
Binwidth = Breaks[2] - Breaks[1]
ggplot(mydata, aes(x=galaxies)) + 
  geom_histogram(aes(y=..density..),# Histogram with density instead of count on y-axis
                 bins = Bins,
                 binwidth=Binwidth,
                 colour="black", fill="#ffaa80") + labs(title = "ggplot: Gaussian Over Smooth")+
  stat_density(kernel='gaussian', bw = 5000, fill = NA, col = "red") +
  stat_density(kernel='gaussian', bw = 800, fill = NA, col = "#660033")+
  stat_density(kernel='gaussian', bw = 100, fill = NA, col = "Blue")
  

  
  
## Base.R
  
  truehist(galaxies,ymax=0.00035,col="green", main="base R: Gaussian Over various bandwidth")+
    lines(density(galaxies,kernel="gaussian",bw=5000),col="red")+
    lines(density(galaxies,kernel="gaussian",bw=800),col="#660033")+
    lines(density(galaxies,kernel="gaussian",bw=100),col="Blue")
    

  
## GGplot

Bins = 6 
Breaks = seq(5000, 35000, length.out = Bins + 1)
Binwidth = Breaks[2] - Breaks[1]
ggplot(mydata, aes(x=galaxies)) + 
  geom_histogram(aes(y=..density..),# Histogram with density instead of count on y-axis
                 bins = Bins,
                 binwidth=Binwidth,
                 colour="black", fill="#ffaa80") +
  labs(title = "ggplot: Triangular Over Smooth")+
  stat_density(kernel='triangular', bw = 5000, fill = NA, col = "red") +
  stat_density(kernel='triangular', bw = 800, fill = NA, col = "#660033")+
  stat_density(kernel='triangular', bw = 100, fill = NA, col = "Blue")
  

  
  
## Base.R
  
  truehist(galaxies,ymax=0.00035,col="green", main="base R: Triangular Over various bandwidth")+
    lines(density(galaxies,kernel="triangular",bw=5000),col="red")+
    lines(density(galaxies,kernel="triangular",bw=800),col="#660033")+
    lines(density(galaxies,kernel="triangular",bw=100),col="Blue")
    
 
  
# Referance : 
# Cami on a D2l group
# https://statisticsglobe.com/kernel-density-plot-in-base-r
# https://stat.ethz.ch/R-manual/R-devel/library/stats/html/density.html 

# Referance: https://stackoverflow.com/questions/59726811/ggplot-density-plot-for-multipl# e-groups

  
```

## Answer-1(c)

       Here, I have used kernel density of Gaussian function and triangular
       function and have used three different bandwidth that is 5000, 800, 
       and 100. With the highest bandwidth, we couldn't see what is happening
       in the graph. I couldn't see the imprints of clusters. Likewise, when 
       the bandwidth is small, there were too many bumps. We couldn't see what
       is happening in the graph. For the proper fit of the kernel density, I
       repeatedly replaced the value of bandwidth and obtained the graph. 
       
    
    d) What is your conclusion about the possible existence of superclusters of galaxies? How many superclusters (1, 2, 3, ... )? (Hint: the existence of clusters implies the existence of empty spaces between galaxies.)


## Answer- 1(d) 

      From the graphs above, we can find the three clusters. However, there can
      be one more cluster in between the data in the middle. Therefore, I can 
      say that there maybe 3 to 4 clusters.

    
    
    e) Fit a finite mixture model using  the Mclust() function in R (from the mclust library). 
    How many clusters did it find?  Did it find the same number of clusters as your graphical 
    inspection? Report parameter estimates and BIC of the best model.
    
```{r}
library(mclust)
model <- Mclust(mydata$galaxies)
summary(model, parameters = TRUE)

plot(model,what="density")+
title ("Density plot of the finite mixture model")

mclustBIC(galaxies)

# Referance:
# https://www.r-bloggers.com/2017/02/an-intro-to-gaussian-mixture-modeling/
# Brian S. Everitt, Sophia Rabe-Hesketh - Handbook of Statistical Analyses Using Stata,
# Third Edition (English and Mandarin Chinese Edition) (2003, Chapman and Hall CRC)
# page - 330-

```

## Answer-1(e)

     From the Mclust, we find out that there are 4 clusters in the data. 
     Whereas form the density plot, we found there are 3 clusters. Therefore
     we can say that there are 3 to 4 clusters in the Galaxies data.



2. (Ex. 8.2 in HSAUR, modified for clarity) The **birthdeathrates** data from **HSAUR3** gives the
birth and death rates for 69 countries (from Hartigan, 1975).

    a) Produce a scatterplot of the data. Estimate the bivariate density and overlay the corresponding contour plot on the scatterplot.
    
```{r}
library(HSAUR3)
library(KernSmooth)
library(reshape2)
library(dplyr)
data(birthdeathrates, ppackage = HSAUR3)
#head(birthdeathrates)
# Build 2d binned kernal estimation
bd_data <- bkde2D(birthdeathrates, bandwidth = sapply(birthdeathrates, dpik))

# Base R
contour(x = bd_data$x1, y = bd_data$x2, z = bd_data$fhat,
        main = "base R: Countor Scatterplot of Birth_Death_Rates",
        xlab="Birth Rates", 
        ylab="Death Rates",
        xlim =c(0,60), ylim = c(0,35))
points(birthdeathrates, pch=16, col="red")

# Referance : Dr. Saunders slide


# ggplot
ggplot(data=birthdeathrates,aes(birth,death)) + 
  geom_density2d(aes(colour=..level..)) + 
  scale_colour_gradient(low="pink",high="Red") + 
  theme_bw() +
  geom_point() +
   labs(title='ggplot: Countor Scatterplot of Birth Death Rates',
       x='Birth Rate',
       y='Death Rate') +
  scale_x_continuous(limits = c(0,60)) +
  scale_y_continuous(limits = c(0,35))


# Referance:
#https://www.r-statistics.com/2016/07/using-2d-contour-plots-within-ggplot2-to-visualize-relationships-between-three-variables/


```
    
    b) What does the contour plot tell you about the structure of the data?

## Answer 2(b)

       Comparaing data for birth rates from 10 to about 50 with the death rates,
       we can tell that the death rate is relatively slow. Also, twice as many
       people are being born than they are dying (i.e., 2:1 birth to death ratio). 
       
    
    c) Produce a perspective plot (persp() in R, ggplot is not required for this question).
    
```{r}

 persp(x = bd_data$x1, y = bd_data$x2, z = bd_data$fhat,
xlab = "Birth Rate",
ylab = "Death Rate",
zlab = "estimated density",
theta = -35, axes = TRUE, box = TRUE,main = "Perspective plot for birthdeathrates data",
col = "#ff99e6")

# Referance:
# Dr Saunders Slide
```
    
    d) Fit a finite mixture model using the Mclust() function in R (from the mclust library).
    Summarize this model using BIC, classification, uncertainty, and/or density plots.
    
```{r}
mod1 <- Mclust(birthdeathrates)
summary(mod1, parameters = TRUE)
names(mod1)
mod1$parameters

# SUmmary: BIC
summary(mclustBIC(birthdeathrates))
summary(mod1$BIC)
#mclustBIC(birthdeathrates)

BIC.data<- as.data.frame(mod1$BIC[,])
BIC.data$NumComp<-rownames(BIC.data)
melted.BIC<- reshape2::melt(BIC.data, var.ids= "NumComp")

# plotting for the BIC
plot(mod1, what='BIC', main = "Base R: Plot of BIC")
#plot(mclustBIC(birthdeathrates))

ggplot(melted.BIC, aes(x=as.numeric(NumComp), y=value, colour=variable, group=variable))+
  scale_x_continuous("Number of Components")+
  scale_y_continuous("BIC")+
  scale_colour_hue("")+
  geom_point()+
  geom_line()+
  theme_bw() + 
  labs(title = "ggplot: BIC")


# For classification
summary(MclustDR(mod1))
plot(mod1, what = 'classification', main = "Base R: classification of plot")


# For ggplot
birthdeathrates %>% mutate(uncertainty = mod1$uncertainty,
                classification = factor(mod1$classification)) %>% 

 ggplot(aes(birth, death, shape = classification, color = classification)) +
  geom_point() + 
  guides(size = guide_legend(), shape = guide_legend()) + theme_classic() +stat_ellipse(level = 0.5, type = "t") + 
  labs(x = "birth", y = "death", title = "ggplot: Plot of classification")


# Uncertainity
summary(mod1$uncertainty)
plot(mod1, what = 'uncertainty', main = "Base R: uncertainity of plot")

birthdeathrates %>% mutate(uncertainty = mod1$uncertainty,
                classification = factor(mod1$classification)) %>% 
  ggplot(aes(birth, death, size = uncertainty, color = classification)) +
  geom_point() + 
  guides(size = guide_legend(), colour = "legend") + theme_classic() +
  stat_ellipse(level = 0.5, type = "t") + 
  labs(x = "birth", y = "death", title = "ggplot: Uncertainty")


plot(mod1, what = "density", main = "Base R: Density plot")
ggplot(birthdeathrates, aes(x = birth, y = death)) +
  geom_point() +
  geom_density_2d(aes(colour=..level..)) + scale_colour_gradient(low="pink",high="Red") +
  labs ( title = "ggplot: Plot of density")


# Referance: 
# https://stat.uw.edu/sites/default/files/files/reports/2012/tr597.pdf page 35

```
    
    e) Discuss the results in the context of Birth and Death Rates.


## Answer 2(e)

      The results and the graphs shows that there are four clusters, The mean birth 
      of the these clusters are nearly equal to 20,34,44,54. Likewise, the mean death 
      of the clusters are nearly equal to 8.5, 9, 12, and 29.
      
    

3. (Ex. 8.3 in HSAUR, modified for clarity) Fit finite mixtures of normal densities individually for each gender in the **schizophrenia** data set from **HSAUR3**. Do your models support the *sub-type model* described in the R Documentation?

    Quote from the R Documentation: *A sex difference in the age of onset of schizophrenia was noted by Kraepelin (1919). Subsequent epidemiological studies of the disorder have consistently shown an earlier onset in men than in women. One model that has been suggested to explain this observed difference is known as the subtype model which postulates two types of schizophrenia, one characterized by early onset, typical symptoms and poor premorbid competence; and the other by late onset, atypical symptoms and good premorbid competence. The early onset type is assumed to be largely a disorder of men and the late onset largely a disorder of women.* (See ?schizophrenia)

```{r}
data("schizophrenia", package = "HSAUR3")
# head(schizophrenia)
ggplot(data = schizophrenia)+
  stat_density(kernal = "Gaussian", adjust = 1,aes(age,fill = gender))+
    labs(title = 'Density plot (gaussian) of Schizophrenia diagnosis data',
       x = "Diagnosis Age",
       y='Density Estimate') +
    theme(
        strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.background = element_blank()
        )

# Referance:
# https://www.slideshare.net/Sheffield_R_/intro-to-ggplot2-sheffield-r-users-group-feb-2015
 

```

Based on the density plot, we can say that the mean of the age at which the schizophenia starts in 
male is about 20, Likewise, in female it is throughout the life.

```{r}
ggplot(data = schizophrenia)+
  geom_histogram(aes(age,fill = gender))+
    labs(title = 'Density plot (gaussian) of Schizophrenia diagnosis data',
       x = "Diagnosis Age",
       y='Density Estimate') +
  scale_fill_manual( values = c("red","blue"))+
  theme(
        strip.background = element_blank(),
        strip.text.y = element_blank(),
        panel.background = element_blank()
        )

# Referances:
# https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf
# http://www.sthda.com/english/wiki/ggplot2-essentials

```


From the histograms, we see that the maximum numbr of men suffered form the disease around the age
of the 20-35 and less at the age of 60. Whereas, in female the disease suffered from the teens and
it continuous to the lifetime.
we can visualize both together and see the same results discussed above by making this plot below:

```{r}
setMen <- subset(schizophrenia, gender=="male")$age
setWmn <- subset(schizophrenia, gender=="female")$age


hist (schizophrenia$age, xlab="Age", ylab="Density", main="Distribution of Schizophrenia by Age", freq=FALSE, ylim=c(0,.075), border=4)
lines(density(setMen), col=1, )
lines(density(setWmn), col=2)
legend(50, 0.05, legend=c("Female", "Male"),
       col=c(1,2), lty=c(1,1), cex=0.8)

# Referance: 
# https://statisticsglobe.com/kernel-density-plot-in-base-r
```



the schizophrenia data by male and female for fit of model analysis by gender:

```{r}
male <- subset(schizophrenia, gender == "male")
female <- subset(schizophrenia, gender == "female")
mod_male <- Mclust(male$age)
plot(mod_male, what = "BIC")+
title(main='BIC Of schizophrenia for male')

mod_female <- Mclust(female$age)
plot(mod_female, what = "BIC")+
title(main = 'BIC of Schizophenia for female')

cat("Summary for male")
print(summary(mod_male, parameters = TRUE))

cat("Summary for female")
print(summary(mod_female, parameters = TRUE))


# Referance 
# Brian S. Everitt, Sophia Rabe-Hesketh - Handbook of Statistical Analyses Using Stata, Third Edition (English and Mandarin Chinese Edition) (2003, Chapman and Hall CRC) page- 130-131
# https://rpubs.com/kkuipers/529695

```

From the model summary above, we can see that the female model showing data points centered at
about 25 and age 47 of age marks, whereas for males it was at around 20 and 27 years of age 
(i.e., within 20s). The BIC plot shows that the optimal number of cluster for both males and 
females is 2.






